Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-10-25 _tex/index.tex
Built site for gh-pages

Moreover, researchers often use these BTL-generated scores or their
transformations separately, conducting further analyses or hypothesis
testing. The CJ literature shows how these scores have been employed to
identify `misfit' judges and stimuli
\citep{Pollitt_2012b, vanDaal_et_al_2017, Goossens_et_al_2018}, detect
biases in judges' ratings \citep{Pollitt_et_al_2003, Pollitt_2012b},
calculate correlations with other scoring methods
\citep{Goossens_et_al_2018, Bouwer_et_al_2023}, or test hypotheses
related to the trait of interest
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.

However, the statistical literature cautions against using estimated
scores for separate analyses and tests. A key consideration is that
BTL-generated scores are parameter estimates that inherently carry
uncertainty. Ignoring this uncertainty when conducting separate analyses
and tests can inflate their precision and statistical power. This, in
turn, increases the risk of committing a type I error
\citep{McElreath_2020}, which is when a null hypothesis is incorrectly
rejected \citep{Everitt_et_al_2010}. To mitigate this risk, principles
from Structural Equation Modeling (SEM)
\citep{Hoyle_et_al_2023, Kline_et_al_2023} and Item Response Theory
(IRT) \citep{deAyala_2009, Fox_2010, vanderLinden_et_al_2017} recommend
conducting these analyses and tests within a structural model that
accounts for both the scores and their uncertainties, rather than
treating them separately. Thus, an integrated approach combining CJ's
structural and measurement models can offer significant advantages.

@@

>>>>

Moreover, researchers often use these BTL-generated scores or their
transformations separately, conducting further analyses or hypothesis
testing. The literature shows how these scores have been employed to
identify `misfit' judges and stimuli
\citep{Pollitt_2012b, vanDaal_et_al_2017, Goossens_et_al_2018}, detect
biases in judges' ratings \citep{Pollitt_et_al_2003, Pollitt_2012b},
calculate correlations with other scoring methods
\citep{Goossens_et_al_2018, Bouwer_et_al_2023}, or test hypotheses
related to the trait of interest
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.

However, the statistical literature cautions against using estimated
scores for separate analyses and tests. A key consideration is that
BTL-generated scores are parameter estimates that inherently carry
uncertainty. Ignoring this uncertainty when conducting separate analyses
and tests can inflate their precision and statistical power, increasing
the risk of committing a type I error \citep{McElreath_2020}, which is
when a null hypothesis is incorrectly rejected
\citep{Everitt_et_al_2010}. To mitigate this risk, principles from
Structural Equation Modeling (SEM)
\citep{Hoyle_et_al_2023, Kline_et_al_2023} and Item Response Theory
(IRT) \citep{deAyala_2009, Fox_2010, vanderLinden_et_al_2017} recommend
conducting these analyses and tests within a structural model that
accounts for both the scores and their uncertainties, rather than
treating them separately. Thus, an integrated approach combining CJ's
structural and measurement models can offer significant advantages.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the BTL model is the statistical articulation of Case 5
\end{enumerate}

From early on in the literature, it has been clear that the BTL model
represents a statistical articulation of Thurstone's Case 5
\citeyearpar{Thurstone_1927}. Talk about \citet{Pollitt_et_al_2003} and
\citet{Bramley_2008}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  But this model hold the greater number of assumptions
\end{enumerate}

Although Case 5 was originally articulated to produce a ``rather coarse
scaling'' of traits \citep[p.~269]{Thurstone_1927}, its assumptions have
become predominant in the literature, but Thurstone's Case 5 is the case
that holds the greater amount of simplifying assumptions.

\citet{Thurstone_1927} justify the use of case V, on multiple
assumptions, but the most important for our purpose are three:

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\tightlist
\item
  related to Case 3, it assumes the correlation between stimuli is zero,
  this translates into the cancellation of judges effects by mean of
  opposing and equally weigthed `mood' and `simultaneous contrasts'
  effects. This is demonstrated using the additivive nature of the logit
  scale, which helps to cancel `bias' judges effects.
\end{enumerate}

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  related to Case 3, it assumes the correlation between stimuli is zero,
  this also translates into the idea that stimuli are the main focus of
  estimation and analysis, but what happens when the focus of analysis
  is the individuals that generated those stimuli. Meaning there is an
  amount of correlation that it is not accounted for.
\end{enumerate}

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  related to case 5, discriminal dispersions of the stimuli are equal,
  but it is not hard to imagine that certain individuals can produce
  good quality texts with more precision than other individuals.
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  What problems does this bring?
\end{enumerate}

This leaves issues such as the presence of judge' biases hinted by
\citet{Bramley_2008} and \citet{Kelly_et_al_2022}, and evidenced by
\citet{Pollitt_et_al_2003}, \citet{Arlett_2003}, and
\citet{Guthrie_2003}

{Reliability is a necessary but not sufficient condition for validity.
Reliability can exist without validity but validity cannot exist without
reliability \citep{Perron_et_al_2015}}.
